Unbalanced Mallows Models for Optimizing Expensive Black-Box Permutation Problems

### 2021-02-02 paper/main.tex
Fix matplotlib fonts in python notebooks

\begin{abstract}
%The abstract should briefly summarize the contents of the paper in
%150--250 words.
  Black-box combinatorial optimization problems arise in practice when the
  objective function is evaluated by means of a simulator or a real-world
  experiment. In such cases, classical techniques such as mixed-integer
  programming and local search cannot be applied. Moreover, each fitness
  evaluation is often expensive in terms of time or resources, thus only a
  limited number of evaluations is possible, typically several orders of
  magnitude smaller than in white-box optimization problems. In the continuous
  case, Bayesian optimization, in particular using Gaussian processes, has
  proven very effective under these conditions. Much less research is available
  in the combinatorial case. In this paper, we propose and analyze UMM, an
  estimation-of-distribution (EDA) algorithm based on a Mallows probabilistic
  model and unbalanced rank aggregation (uBorda). Experimental results on
  black-box versions of LOP, QAP and PFSP show that UMM is able to match, and
  sometimes surpass, the solutions obtained by CEGO, a Bayesian optimization
  algorithm for combinatorial optimization. %, specially when the budget of solutions evaluated approaches 400. 
  Moreover, the computational complexity of UMM increases linearly with both the number of function evaluations and the permutation size.
%  an order of magnitude smaller than CEGO, which makes it specially useful when
  %both computation time and number of evaluations are of concern.
  \sloppy
  %\textcolor{red}{16 pages max, deadline: 1 November 2020}
  \keywords{Combinatorial optimization \and Bayesian optimization \and
    Expensive black-box optimization \and Estimation of distribution
    algorithms}
\end{abstract}

@@

In many practical optimization problems, the objective function is not
explicitly available and solutions are evaluated by means of expensive
prediction models, simulations or physical experiments. When decision variables
are continuous, the use of Bayesian surrogate-models (e.g., Gaussian processes)
is widespread in
optimization~\citep{JonSchWel98go,ForKea2009surrogate}. Motivated by this
success, there have been attempts at adapting such surrogate-based optimization
algorithms to the combinatorial case~\cite{MorKat2011:evo}, a notable example being Combinatorial
Efficient Global Optimization
(CEGO)~\citep{ZaeStoBar2014:ppsn,ZaeStoFriFisNauBar2014}. However, the
ruggedness of combinatorial landscapes, which makes local search particularly
effective in the white-box context, lessens the effectiveness of global
surrogate models~\citep{EriPeaGar2019scalable}. Moreover, surrogate models are
expensive to train and optimize. The additional time required by the Bayesian
optimizer may impose a significant overhead in computation time, specially if
computation time of each function evaluation is measured in ``few'' minutes or
hours rather than days and when ``expensive'' refers to resources or economical
cost rather than wall-clock time. \citet{PerLopStu2015si} recently showed that
ant colony optimization (ACO) %~\citep{DorStu2004:book}
is competitive with CEGO
on a black-box version of the travelling salesman problem under a budget of
$1\,000$ function evaluations.  On the other hand, $1\,000$ evaluations is
still a relatively large budget for CEGO, whereas ACO was not designed for such
short budgets.
%Finally, we note that learning Budget-constrained Combinatorial is a hot topic for a variety of structures besides permutation~\cite{manchanda2020}. 
\manuel{- the authors assigned to black-box optimization problems characteristics that they do not own. For instance they say that "local search cannot be used".
 they mixed expensive to evaluate functions with black-box functions and of course they are different. In fact much of the metaheuristics do not mind
 if the problem is withe-box or black-box as they do not use static information from the function in the optimization process. In this case I understand
 that they have a black-box problem where the objective function is costly to evaluate. This should be clear and I would recommend to delete other comments
 related with white-box and black-box.
}{}

\ekhine{6. When the main contribution of the paper is discussed it is stated that this contribution is the Unbalanced Mallows *model* (which does not seem to be true as the models used in the paper seem to be known already). Right after that it is stated that the contribution is *a novel algorithm* which, I think, reflects the contents of the paper much better.
}{}
In this work, we propose and analyze Unbalanced Mallows Model (UMM), a
population-based probabilistic algorithm specifically designed for optimizing
black-box problems over a permutation landscape under a limited budget. There
exists a wide body of research on probabilistic models for permutations, such
as Mallows~\cite{FliVer1986}, that have already been been used in optimization
problems successfully, with the most prominent examples being estimation of
distribution algorithms
(EDAs)~\cite{CebIruMen2014eda}. ACO may also be considered
a type of EDA, since it builds a probability distribution model from which
solutions are sampled with a bias towards the best solutions evaluated so
far. These classical EDAs  assume that function
evaluations are cheap and/or the fitness function is white-box.%\MANUEL{Do EDAs really assume that the fitness function is white-box? ACO doesn't: They are hybridized with local search }
%The proposed UMM drastically differs from the above algorithms.

>>>>

\begin{abstract}
%The abstract should briefly summarize the contents of the paper in
%150--250 words.
  Expensive black-box combinatorial optimization problems arise in practice
  when the objective function is evaluated by means of a simulator or a
  real-world experiment. Since each fitness evaluation is expensive in terms of
  time or resources, only a limited number of evaluations is possible,
  typically several orders of magnitude smaller than in non-expensive problems. In
  this scenario, classical optimization methods such as mixed-integer
  programming and local search are not useful.  In the continuous case,
  Bayesian optimization, in particular using Gaussian processes, has proven
  very effective under these conditions. Much less research is available in the
  combinatorial case. In this paper, we propose and analyze UMM, an
  estimation-of-distribution (EDA) algorithm based on a Mallows probabilistic
  model and unbalanced rank aggregation (uBorda). Experimental results on
  black-box versions of LOP, QAP and PFSP show that UMM is able to match, and
  sometimes surpass, the solutions obtained by CEGO, a Bayesian optimization
  algorithm for combinatorial
  optimization. %, specially when the budget of solutions evaluated approaches 400.
  Moreover, the computational complexity of UMM increases linearly with both
  the number of function evaluations and the permutation size.
%  an order of magnitude smaller than CEGO, which makes it specially useful when
  %both computation time and number of evaluations are of concern.
  \sloppy
  %\textcolor{red}{16 pages max, deadline: 1 November 2020}
  \keywords{Combinatorial optimization \and Bayesian optimization \and
    Expensive black-box optimization \and Estimation of distribution
    algorithms}
\end{abstract}

@@

In many practical optimization problems, the objective function is not
explicitly available and solutions are evaluated by means of expensive
prediction models, simulations or physical experiments. When decision variables
are continuous, the use of Bayesian surrogate-models (e.g., Gaussian processes)
is widespread in
optimization~\citep{JonSchWel98go,ForKea2009surrogate}. Motivated by this
success, there have been attempts at adapting such surrogate-based optimization
algorithms to the combinatorial case~\cite{MorKat2011:evo}, a notable example being Combinatorial
Efficient Global Optimization
(CEGO)~\citep{ZaeStoBar2014:ppsn,ZaeStoFriFisNauBar2014}. However, the
ruggedness of combinatorial landscapes, which makes local search particularly
effective for combinatorial problems, lessens the effectiveness of global
surrogate models~\citep{EriPeaGar2019scalable}. Moreover, surrogate models are
expensive to train and optimize. The additional time required by the Bayesian
optimizer may impose a significant overhead in computation time, specially if
computation time of each function evaluation is measured in ``few'' minutes or
hours rather than days and when ``expensive'' refers to resources or economical
cost rather than wall-clock time. \citet{PerLopStu2015si} recently showed that
ant colony optimization (ACO) %~\citep{DorStu2004:book}
is competitive with CEGO
on a black-box version of the travelling salesman problem under a budget of
$1\,000$ function evaluations.  On the other hand, $1\,000$ evaluations is
still a relatively large budget for CEGO, whereas ACO was not designed for such
short budgets.
%Finally, we note that learning Budget-constrained Combinatorial is a hot topic for a variety of structures besides permutation~\cite{manchanda2020}. 
\MANUEL{- the authors assigned to black-box optimization problems characteristics that they do not own. For instance they say that "local search cannot be used".
 they mixed expensive to evaluate functions with black-box functions and of course they are different. In fact much of the metaheuristics do not mind
 if the problem is withe-box or black-box as they do not use static information from the function in the optimization process. In this case I understand
 that they have a black-box problem where the objective function is costly to evaluate. This should be clear and I would recommend to delete other comments
 related with white-box and black-box. MANUEL: FIXED
}

\EKHINE{6. When the main contribution of the paper is discussed it is stated that this contribution is the Unbalanced Mallows *model* (which does not seem to be true as the models used in the paper seem to be known already). Right after that it is stated that the contribution is *a novel algorithm* which, I think, reflects the contents of the paper much better.
}
In this work, we propose and analyze Unbalanced Mallows Model (UMM)\MANUEL{Perhaps we need to rename the algorithm to UMM Algorithm or UMM Optimizer to avoid the criticism of the reviewer that we do not propose the model?}, a
population-based probabilistic algorithm specifically designed for optimizing
black-box problems over a permutation landscape under a limited budget. There
exists a wide body of research on probabilistic models for permutations, such
as Mallows~\cite{FliVer1986}, that have already been  used in optimization
problems successfully, with the most prominent examples being estimation of
distribution algorithms
(EDAs)~\cite{CebIruMen2014eda}. ACO may also be considered
a type of EDA, since it builds a probability distribution model from which
solutions are sampled with a bias towards the best solutions evaluated so
far. These classical EDAs  assume that function
evaluations are cheap and/or the fitness function is white-box.%\MANUEL{Do EDAs really assume that the fitness function is white-box? ACO doesn't: They are hybridized with local search }
%The proposed UMM drastically differs from the above algorithms.
